---
title: "Exam 2"
author: "Brianna Baker"
date: "6/26/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exam 2

Load the library and save the data frame as inequality_data.


```{r}
library(rio)
education_data = import("~/Desktop/inequality.xlsx", which =1)
#saving data frame
inequality_data <- education_data
#removing education_data from environment
rm(education_data)
```

## Question 3

This is a cross-sectional dataset because it provides a snapshot of data from the same time and not change over time. We can see this is the code below. 

```{r}
summary(inequality_data$year)
```


##Using the subset command to show the inequality_gini scores for Denmark and Sweden
```{r}
library(dplyr)
subset(inequality_data, country == "Denmark")
subset(inequality_data, country == "Sweden")
```
##Inequality score for Brazil. 
```{r}
subset(inequality_data, country == "Brazil")
```

##Question 6
Since Denamrk and Sweden were described as having "optimal Gini index scores," and they have much lower scores than Brazil, it appears that it is better to have a low inequality gini score. 

###Quick peak at data frame:
```{r}
head(inequality_data)
```
###Removing accent with new function "accent.remove."
```{r}
#change default text encoding to UTF-8
#define a function
remove.accents <- function(s) {
  #1 character substiutions
  old1 <- "ú"
  new1 <- "u"
  s1 <- chartr(old1, new1, s)
}
#remove accents
inequality_data$country <- remove.accents(inequality_data$country)
```

###Quick peak to show accent removal
```{r}
head(inequality_data)
```

###Sorting data by countries with lowest inequality_gini scores. 
```{r}
inequality_data <- inequality_data[order(inequality_data$inequality_gini),]
#top 5 countries
head(inequality_data)
```

###Mean of inequality gini scores
```{r}
mean(inequality_data$inequality_gini, na.rm = TRUE)
```
###Using if else to recode variables and assign values based on relation to mean. 
for (r in 1:nrow(inequality_data)){
  for(c in 1:ncol(inequality_data)) {
    if(inequality_data$inequality_gini[r,c] > 36.81375) {
      inequality_data[r,c] = "high_inequality"
    }
    else if(inequality_data$inequality_gini[r,c] < 36.81375) {
      inequality_data[r,c] = "low_inequality"
    }
    else{
    }
  }
}

### Question 13
```{r}
#create vector
actors <- c('World Bank', 'African Development Bank', 'Bill and Melinda Gates Foundation')

#create for statement
for (i in actors) {
  print (i)
}

```
###Question 14
I chose the variable "Employment to population ratio" to demonstrate inequality because the comparison could show whether or not the majority of a country has employed inhabitants or not. The more employed, the lower the inquality- is my prediction. 

```{r}
library(devtools)
library(remote)
#add some data from the World Development Indicators (WDI)
library(WDI)
employment_ratio = WDI(country = "all",
                      indicator = "SL.EMP.TOTL.SP.ZS", 
                      start = 2015, end = 2015, extra = FALSE, cache = NULL)
#quick peak 
summary(employment_ratio)

#changing name of variable 
library(data.table)
#changing name of column to something easier to interpret
setnames(employment_ratio, "SL.EMP.TOTL.SP.ZS", "employment_ratio")
```

###Merge new variable into other dataset
```{r}
merged_df = left_join(inequality_data,
                        employment_ratio,
                        by = c("iso2c", "year"))

```
#drop country.y and rename country.x as country
library(tidyverse)
merged_df <-
  merged_df %>%
  select(-c("country.x")) %>%
  rename("country" = "country.y")
  
###Removing NAs
```{r}
na.omit(merged_df, select =c("employment_ratio", "inequality_gini"))
```

##Filtering out data with inequality gini scores greater than 30
```{r}
data_greater_30 <- 
  merged_df %>% 
  dplyr::filter(inequality_gini > 30)
```
###Count how many countries contain "ai"
```{r}
grep("ai", data_greater_30)
```

###Using lapply to take sum of inequality gini
```{r}
lapply(data_greater_30$inequality_gini, sum)
```
###Labeling variables and save as Stata
library(labelled)
#use `` for variable names and "" for labels
var_label(merged_df) <- list(`country` = "country"
                               `year` = "year",
                               `inequality_gini` = "inequality gini score",
                               `population` = "population (inhabitants)",
                               `iso2c` = "ISO-2 country code",
                               `employment_ratio` = "ratio of employment to population")
```{r}

                              
#save the data frame as a Stata dataset
library(rio)
#export(merged_df, "final_data.dta")

```


